---
title: "Homework 5"
author: "Thomas O'Leary"
output:
rmarkdown::html_document:
    theme: lumen
    highlight: github
---

## Working with matrices, lists, and data frames

### 1. Assign to the variable `n_dims` a single random integer between 3 and 10.

```{r}
n_dims <- sample(3:10, size = 1)
```

#### Create a vector of consecutive integers from 1 to $\mbox{n_dims}^2$.

```{r}
vec <- seq(from = 1, to = n_dims^2)
```


#### Use the sample function to randomly reshuffle these values.

```{r}
vec_shuf <- sample(vec, size = length(vec))
```

#### create a square matrix with these elements.

```{r}
mat <- matrix(vec_shuf, nrow = n_dims)
```

#### print out the matrix.

```{r}
mat
```

#### find a function in r to transpose the matrix.

```{r}
t_mat <- t(mat)
```

#### print it out again and note how it has changed.

```{r}
t_mat
```

#### calculate the sum and the mean of the elements in the first row and the last row.

```{r}
# sum of first row
sum(t_mat[1,])
#sum of last row
sum(t_mat[nrow(t_mat),])
# mean of first row
mean(t_mat[1,])
# mean of last row
mean(t_mat[nrow(t_mat),])
```

#### read about the `eigen()` function and use it on your matrix

```{r}
t_mat_eigen <- eigen(t_mat)
t_mat_eigen
```

#### look carefully at the elements of `$values` and `$vectors`. What kind of numbers are these?

```{r}
# eigen values
t_mat_eigen$values
# eigen vectors
t_mat_eigen$vectors
```


#### dig in with the `typeof()` function to figure out their type.

```{r}
# these are complex numbers

# t_mat_eigen$values is a vector of length nrow(t_mat)
typeof(t_mat_eigen$values)

# t_mat_eigen$vectors is a matrix of the same dimensions as t_mat
typeof(t_mat_eigen$vectors)
```

#### $\checkmark$ if have set your code up properly, you should be able to re-run it and create a matrix of different size because `n_dims` will change.

### 2. Create a list with the following named elements:

#### `my_matrix`, which is a 4 x 4 matrix filled with random uniform values

```{r}
my_matrix <- matrix(runif(16), nrow = 4)
```

####  `my_logical` which is a 100-element vector of TRUE or FALSE values. Do this efficiently by setting up a vector of random values and then applying an inequality to it.

```{r}
my_logical <- runif(100) < 0.5
```

#### `my_letters`, which is a 26-element vector of all the lower-case letters in random order.

```{r}
my_letters <- sample(letters, size = 26)
```

#### `my_list` which is all those elements compiled into a list

```{r}
my_list <- list(my_matrix = my_matrix,
                my_logical = my_logical,
                my_letters = my_letters)

my_list
```


#### create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.

```{r}
new_list <- list(my_list$my_matrix[2,2], 
                 my_list$my_logical[2],
                 my_list$my_letters[2])
new_list
```

#### use the `typeof()` function to confirm the underlying data types of each component in this list

```{r}
# double
typeof(new_list[[1]])
# logical
typeof(new_list[[2]])
# character
typeof(new_list[[3]])
```

#### combine the underlying elements from the new list into a single atomic vector with the `c()` function.

```{r}
atomic_vec <- c(new_list[[1]], 
                new_list[[2]], 
                new_list[[3]])
atomic_vec
```

#### what is the data type of this vector?

```{r}
typeof(atomic_vec)
```

### 3. Create a data frame with two variables (= columns) and 26 cases (= rows).

#### call the first variable `my_unis` and fill it with a random uniform value from 0 to 10

#### call the second variable `my_letters` and fill it with 26 capital letters in random order.

```{r}
df <- data.frame(my_unis = runif(n = 26, min = 0, max = 10),
                 my_letters = sample(LETTERS, size = 26))
df
```

#### for the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with `NA`.

```{r}
df$my_unis[sample(1:26, size = 4)] <- NA
```


#### for the first variable, write a single line of R code to identify which rows have the missing values.

```{r}
my_unis_na <- which(is.na(df$my_unis))
my_unis_na
```

#### for the second variable, sort it in alphabetical order

```{r}
sort(df$my_letters)
```

#### calculate the column mean for the first variable.

```{r}
mean(df$my_unis, na.rm = TRUE)
# alternatively, although I think this is less readable
mean(df$my_unis[-my_unis_na])
```

