---
title: "Homework 4"
author: "Thomas O'Leary"
output:
rmarkdown::html_document:
    theme: lumen
    highlight: github
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = TRUE)
```

# 1. Suppose *x* = 1.1, *a* = 2.2, and *b* = 3.3. Assign each expression to the value of the variable *z*and print the value stored in *z*.
    
```{r}
x <- 1.1
a <- 2.2
b <- 3.3
```

##    a) $x^{a^b}$
    
```{r}
z <- x^a^b 
z
```

##    b) $(x^{a})^b$

```{r}
z <- (x^a)^b 
z
```
##    c) $3x^3 + 2x^2 + 1$

```{r}
z <- 3*(x^3) + 2*(x^2) + 1
z
```    


# 2. Using the `rep` and `seq` functions, create the following vectors:

##    a) $(1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)$
    
```{r}
c(seq(from = 1, to = 8), seq(from = 7, to = 1))
```
    
##    b) $(1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)$
    
```{r}
rep(seq(1:5), times = c(seq(1:5)))
```    

##    c) $(5,4,4,3,3,3,2,2,2,2,1,1,1,1,1)$
    
```{r}
rep(seq(from = 5, to = 1), times = c(seq(1:5)))
```       
    
    
# 3. Create a vector of two random uniform numbers. In a spatial map, these can be interpreted as *x* and *y* coordinates that give the location of an individual (such as a marked forest tree in a plot that has been mapped). Using one of R's inverse trigonometry functions (`asin()`, `acos()`, or `atan()`), convert these numbers into polar coordinates (If you don't know what polar coordinates are, read about them on the web or in your calculus textbook).

```{r}
set.seed(1)
v <- runif(2)
# polar coordinates: (r, theta)
# r = sqrt(x^2 + y^2)
# theta = atan(y/x)

r <- sqrt(v[1] + v[2]) 
theta <- atan(v[2]/v[1])

paste("r =", round(r, 4), "; theta =", round(theta, 4))
```  

# 4. Suppose that `queue <- c("sheep", "fox", "owl", "ant")` and that `queue` represents the animals that are lined up to enter Noah's Ark, with the sheep at the front of the line. Using R expressions, update the queue successively as 

```{r}
queue <- c("sheep", "fox", "owl", "ant")
queue
```

##    a) the serpent arrives;
```{r}
queue <- c(queue, "serpent")
queue
```    
    
##    b) the sheep enters the ark;
    
```{r}
queue <- queue[-1]
queue
```    
##    c) the donkey arrives and talks his way to the front of the line;
    
```{r}
queue <- c("donkey", queue)
queue
```    

##    d) the serpent gets impatient and leaves;
    
```{r}
queue <- queue[-which(queue == "serpent")]
queue
``` 

##    e) the owl gets bored and leaves;
    
```{r}
queue <- queue[-which(queue == "owl")]
queue
```     
    
##    f) the aphid arrives and the ant invites him to cut in line.
    
```{r}
pos_ant <- which(queue == "ant")
queue <- c(queue[1:(pos_ant - 1)], "aphid", queue[pos_ant:length(queue)])
queue
```    
    
##    g) Finally, determine the position of the aphid in the line.
    
```{r}
pos_aphid <- which(queue == "aphid") 
pos_aphid
```    

# 5. Use R to create a vector of all of the integers from 1 to 100 that are not divisible by 2, 3, or 7.


```{r}
v <- which(1:100 %% 2 != 0 & 1:100 %% 3 != 0 & 1:100 %% 7 != 0)
v
```   
