---
title: "Homework 8"
author: "Thomas O'Leary"
output:
  rmarkdown::html_document:
    theme: lumen
---

## Data manipulations using the `dplyr` package  

This homework assignment focus on data manipulation in `R`. Complete these problems using the `dplyr` package, and if you have time, try each of these exercises using the subsetting methods in base `R` that have been covered in lecture. 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

require(tidyverse)
```

1. $\checkmark$ Examine the structure of the iris dataset. How many observations and variables are in the dataset? 

```{r}
str(iris)
```

2. $\checkmark$ Create a new data frame `iris1` that contains only the species *virginica* and *versicolor* with sepal lengths longer than 6 cm and sepal widths longer than 2.5 cm. How many observations and variables are in the dataset?

```{r}
sp <- c("virginica", "versicolor")

iris1 <- iris %>%
  filter(Species %in% c("virginica", "versicolor") &
           Sepal.Length > 6 &
           Sepal.Width > 2.5)

str(iris1)
```

3. $\checkmark$ Now, create a `iris2` data frame from `iris1` that contains only the columns for Species, Sepal.Length, and Sepal.Width. How many observations and variables are in the dataset?  

```{r}
iris2 <- iris1 %>%
  select(Species, Sepal.Length, Sepal.Width)

str(iris2)
```

4. $\checkmark$ Create an `iris3` data frame from `iris2` that orders the observations from largest to smallest sepal length. Show the first 6 rows of this dataset.   

```{r}
iris3 <- iris2 %>%
  arrange(desc(Sepal.Length))

head(iris3)
```

5. $\checkmark$ Create an `iris4` data frame from `iris3` that creates a column with a sepal area (length * width) value for each observation. How many observations and variables are in the dataset?

```{r}
iris4 <- iris3 %>%
  mutate(Sepal.Area = Sepal.Length * Sepal.Width)

head(iris4)
```

6. $\checkmark$ Create `iris5` that calculates the average sepal length, the average sepal width, and the sample size of the entire `iris4` data frame and print `iris5`.


```{r}
iris5 <- iris4 %>%
  summarize(Mean.Sepal.Width = mean(Sepal.Width),
            Mean.Sepal.Length = mean(Sepal.Length),
            Sample.Size = n())

iris5
```


7. $\checkmark$ Finally, create `iris6` that calculates the average sepal length, the average sepal width, and the sample size *for each species* of in the `iris4` data frame and print `iris6`.

```{r}
iris6 <- iris4 %>%
  group_by(Species) %>%
  summarize(Mean.Sepal.Width = mean(Sepal.Width),
            Mean.Sepal.Length = mean(Sepal.Length),
            Sample.Size = n())

iris6
```

8. $\checkmark$ In these exercises, you have successively modified different versions of the data frame `iris1 iris1 iris3 iris4 iris5 iris6`. At each stage, the output data frame from one operation serves as the input to the next. 

A more efficient way to do this is to use the pipe operator `%>%` from the `tidyr` package. See if you can rework all of your previous statements into an extended piping operation that uses `iris` as the input and generates `iris6` as the output.


```{r}
iris6 <- iris %>%
  filter(Species %in% c("virginica", "versicolor") &
           Sepal.Length > 6 &
           Sepal.Width > 2.5) %>%
  select(Species, Sepal.Length, Sepal.Width) %>%
  arrange(desc(Sepal.Length)) %>%
  mutate(Sepal.Area = Sepal.Length * Sepal.Width) %>%
  group_by(Species) %>%
  summarize(Mean.Sepal.Width = mean(Sepal.Width),
            Mean.Sepal.Length = mean(Sepal.Length),
            Sample.Size = n())
  
iris6
```

