---
title: "Homeworks 11 & 12"
subtitle: "Batch processing & ggplot"
author: "Thomas O'Leary"
output:
  rmarkdown::html_document:
    theme: lumen
    toc: true
    toc_float: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

# Homework 11: Batch processing

Repeat the exercise from the Batch Processing Lecture (7th April), but do it using real data sets rather than purely simulated. Check with folks in your lab to see if there are multiple data sets available for analysis. Stick to simple data analyses and graphics, but try to set it up as a batch process that will work on multiple files and save summary results to a common file.

For this homework, I am taking spec data from **Gretchen Thompson**'s honors thesis measuring ADH activity in different populations of _D. melanogaster_, where she was trying to measure K<sub>m</sub>'s of these enzymes at different temperatures. This type of data and analysis is prime fodder for batch processing as the spec spits out many absorbance values for a bunch of samples at a few different time points. Then, velocity needs to be calculated based on a bunch of linear regressions. Then, a Michaelis–Menten curve needs be fitted on the velocities at each concentration, to find K<sub>m</sub> values. 


## Functions

```{r}
# ------------------------------------------------------------------------------
# Batch processing spec data
# April 16, 2020
# TS O'Leary
# ------------------------------------------------------------------------------


# ------------------------------------------------------------------------------
# Function: read_tidy_spec_csv
# Description: uses read_csv and tidys the data
# Inputs: .csv file 
# Outputs: data.frame

require(tidyverse)

read_tidy_spec_csv <- function(file) {
  
  read_csv(file) %>%
    pivot_longer(-c("cycle", "time", "temp"), 
                 names_to = "conc", 
                 values_to = "abs") %>%
    mutate(conc = as.numeric(str_replace_all(conc, "_.", ""))) %>%
    mutate(file = file,
           temp_group = case_when(temp > 17.5 & temp < 19.5 ~ "18.5",
                                  temp > 22.5 & temp < 24.5 ~ "23.5",
                                  temp > 27.5 & temp < 29.5 ~ "28.5",
                                  temp > 32.5 & temp < 34.5 ~ "33.5",
                                  temp > 37.5 & temp < 39.5 ~ "38.5"))
  
} 
# End function -----------------------------------------------------------------


# ------------------------------------------------------------------------------
# Function: read_bind_tidy_csv
# Description: Read all the spec csvs in a directory
# Inputs: dir is the directory with csv files to load and bind
# Outputs: large binded tibble

require(tidyverse)

read_bind_tidy_csv <- function(dir) {
  setwd(dir)
  files <- list.files()
  dfs <- vector(mode = "list", length = length(files))
  
    for (i in seq_along(files)){
      dfs[[i]] <- read_tidy_spec_csv(files[i])
    }
  
  bind_rows(dfs)
} 
# End function -----------------------------------------------------------------


# ------------------------------------------------------------------------------
# Function: calc_velocity
# Description: Linear regression to determine velocity of the reaction in 
#              absorbance over time at each conc for each temp_group
# Inputs: tibble from read_bind_tidy_csv() with conc, temp_group, 
#         abs, & time columns
# Outputs: tibble with velocity estimates

require(tidyverse)
require(broom)

calc_velocity <- function(data) {
  data %>%
    group_by(conc, temp_group) %>%
    nest() %>%
    mutate(
      fit = map(data, ~ lm(abs ~ time, data = .x)),
      tidied = map(fit, tidy)
    ) %>% 
    unnest(tidied) %>%
    filter(term == "time" & conc != "0" & estimate > 0) %>%
    select(temp_group, conc, estimate) %>%
    rename(velocity = estimate)
} 
# End function -----------------------------------------------------------------


# ------------------------------------------------------------------------------
# Function: calc_Km_Vm
# Description: Calculate Vmax and Km for Michaelis–Menten curve
# Inputs: tibble with velocity estimates for each conc and temp_group
# Outputs: tibble with estimates for Vmax and Km

require(tidyverse)
require(broom)

calc_Km_Vm <- function(data) {
  data %>%
    filter(conc != 0 & velocity > 0) %>%
    group_by(temp_group) %>%
    nest() %>%
    mutate(
      fit = map(data, ~ nls(formula = 
                              velocity ~ SSmicmen(conc, Vmax, Km), data = .), 
                data = .x),
      tidied = map(fit, tidy)
    ) %>%
    unnest(tidied)
} 
# End function -----------------------------------------------------------------
```

## Program body

```{r}

# Load packages
require(tidyverse)
require(broom)
require(RColorBrewer)

# Calculate reaction velocities at each [EtOH] and each Temp group -------------
# Determine the Km and Vmax from the fit of the Michaelis–Menten curves --------

# ADH-fast enzyme
data_f <- read_bind_tidy_csv(dir = here::here("Homework/spec_data/adh-f"))
head(data_f)
data_velo_f <- calc_velocity(data_f)
data_velo_f
data_Km_Vm_f <- calc_Km_Vm(data_velo_f)
data_Km_Vm_f

# ADH-slow enzyme
data_s <- read_bind_tidy_csv(dir = here::here("Homework/spec_data/adh-s")) %>%
  filter(str_detect(file, "deg"))
data_velo_s <- calc_velocity(data_s)
data_Km_Vm_s <- calc_Km_Vm(data_velo_s)

# VTECK8 ADH enzyme
data_vt <- read_bind_tidy_csv(dir = here::here("Homework/spec_data/vt")) %>%
  filter(str_detect(file, "deg"))
data_velo_vt <- calc_velocity(data_vt)
data_Km_Vm_vt <- calc_Km_Vm(data_velo_vt)

# Chiapas ADH enzyme
data_chiapas <- read_bind_tidy_csv(dir = here::here("Homework/spec_data/chiapas")) %>%
  filter(str_detect(file, "deg") & !is.na(abs))
data_velo_chiapas <- calc_velocity(data_chiapas)
data_Km_Vm_chiapas <- calc_Km_Vm(data_velo_chiapas)
```

# Homework 12: Advanced ggplotting

For this exerise, use your newly-developed ggplot chops to create some nice graphs from your own data (If you do not have a good data frame to use for graphics, use one of the many built-in data frames from R (other than mpg, which we are using in class)). Experiment with different themes, theme base sizes, aesthetics, mappings, and faceting. When you are finished, try exporting them to high quality pdfs, jpgs, eps files, or other formats that you would use for submission to a journal.

## Functions

```{r}
# Plotting functions

# ------------------------------------------------------------------------------
# Function: plot_mm_curve
# Description: Plot the Michaelis-Menten curves
# Inputs: tibble with conc and velocity columns
# Outputs: plot

require(tidyverse)

plot_mm_curve <- function(data, 
                          palette = c("skyblue", "blue", "purple", "red", "darkred")) {
  
  ggplot(data, aes(x = conc, y = velocity, fill = temp_group)) +
    geom_point(pch = 21, size = 3, color = "grey50", alpha = 0.75) +
    geom_smooth(method = "nls", 
                formula = y ~ SSmicmen(x, Vmax, Km),
                data = data,
                se = FALSE,
                aes(color = temp_group), 
                show.legend = FALSE) +
    scale_fill_manual(name = "Temperature",
                      values = palette,
                      labels = c("18.5°C", 
                                 "23.5°C", 
                                 "28.5°C", 
                                 "33.5°C", 
                                 "38.5°C")) +
    scale_color_manual(values = palette) + 
    labs(x = "Ethanol (mM)",
         y = expression("Reaction Velocity (Abs/sec)")) + 
    theme_classic()
} 
# End function -----------------------------------------------------------------


# ------------------------------------------------------------------------------
# Function: plot_km
# Description: Plot km values with std.error bars
# Inputs: tibble with Km estimate and error
# Outputs: plot

require(tidyverse)

plot_km <- function(data, 
                    palette = brewer.pal(4, "Set2"),
                    legend.labels = c( "VT-ECK 8", "Chiapas", "Fast", "Slow")) {
  ggplot(data, aes(x = temp_group, y = estimate_Km, fill = enzyme)) +
    geom_errorbar(aes(ymin = estimate_Km - std.error_Km, 
                      ymax = estimate_Km + std.error_Km), 
                  width = .2,
                  position = position_dodge(0.5)) +
    geom_point(pch = 21, position = position_dodge(0.5), size = 6) +
    labs(x = "Temperature (°C)",
         y = expression("K"[m]*"  (mM EtOH)")) +
    scale_fill_manual(name = "ADH enzyme",
                      values = palette,
                      labels = legend.labels) +
    theme_classic()
} 
# End function -----------------------------------------------------------------
```

## Michaelis–Menten curves

```{r}
# ADH-fast enzyme
plot_mm_curve(data_velo_f)

# ADH-slow enzyme
plot_mm_curve(data_velo_s)

# VTECK8 ADH enzyme
plot_mm_curve(data_velo_vt)

# Chiapas ADH enzyme
plot_mm_curve(data_velo_chiapas)
```

## Plot K<sub>m</sub> values

```{r}
# Combine Km tibbles -----------------------------------------------------------
df_km_plot <- bind_rows("fast" = data_Km_Vm_f, 
                        "slow" = data_Km_Vm_s, 
                        "vt" = data_Km_Vm_vt, 
                        "chiapas" = data_Km_Vm_chiapas, 
                        .id = "enzyme") %>%
  mutate(enzyme = factor(enzyme, levels = c("vt", "chiapas", "fast", "slow"))) %>%
  select(enzyme, temp_group, term, estimate, std.error) %>%
  pivot_wider(names_from = term, 
              values_from = c(estimate, std.error))

df_km_plot


# Plot all Km's on the same chart
plot_km(df_km_plot)

# Plot all just fast and slow enzymes
plot_km(df_km_plot%>%
         filter(enzyme %in% c("fast", "slow" )),
        palette = brewer.pal(4, "Set2")[3:4],
        legend.labels = c("Fast", "Slow"))

# Plot all just vt and chiapas
plot_km(df_km_plot%>%
         filter(enzyme %in% c("vt", "chiapas" )),
        palette = brewer.pal(4, "Set2")[3:4],
        legend.labels = c("VT-ECK 8", "Chiapas"))
```
